In the Crucible: Massive Compact Galaxies with High-Velocity Outflows and Eddington-Limited Star Formation

### 2012-04-20 trunk/tex/sigmasfr/ms.tex
version sent around on fri, apr 20

The central regions of elliptical galaxies are thought to form in
compact starbursts \citep[e.g.,][]{kor09,hop09}.  Feedback associated
with such starbursts can produce outflows driven by thermal energy
from supernova explosions \citep[e.g.,][]{che85}, stellar winds
\citep[e.g.,][]{lei92}, and momentum input from both supernova ram
pressure and radiation pressure on dust grains \citep[e.g.,][]{mur05}.
It has been argued that such feedback imposes a limit on the maximum
star-formation rate surface density ($\sigmasfr$) for starbursts
\citep[e.g.,][]{leh96,meu97,mur05,tho05} and the maximum stellar
surface density for elliptical galaxies and star clusters
\citep[e.g.,][]{hop10}.

@@

In this Letter, we present results for a sample of massive galaxies at
$z\sim0.6$ that exhibit $\gtrsim1000$~\kms\ outflows, expanding on the
sample from \citet{tre07}.  We seek to test whether the energetic
outflows in these galaxies could have been driven by feedback from
starbursts with very large star-formation rate (SFR) surface densities
($\sigmasfr$).  Our analysis combines galaxy sizes obtained with the
Hubble Space Telescope (HST) with star-formation rates and stellar
masses estimated using photometry from the Wide-field Infrared Survey
Explorer (WISE), the Spitzer Space Telescope, the Sloan Digital Sky
Survey (SDSS), and the Galaxy Evolution Explorer (GALEX).

@@

% We gathered photometry from the All-Sky Release of the Wide-field
% Infrared Survey Explorer \citep[WISE,][]{wri10}, the Seventh Data
% Release of the Sloan Digital Sky Survey \citep[SDSS,][]{aba09}, and
% General Release 6 from the Galaxy Evolution Explorer
% \citep[GALEX,][]{mar05galex}.  
We gathered photometry from the WISE All-Sky Release, the SDSS Seventh
Data Release, and GALEX General Release 6.  We also obtained
$5\times30$~sec dithered exposures at 3.6~$\mu$m and 4.5~$\mu$m for
all sources with the Infrared Array Camera on the Spitzer Space
Telescope as part of General Observer program 60145 (PI: C. Tremonti).
We used the post--basic calibrated data to perform aperture photometry
on all sources, and we also used the
APEX\footnote{http://irsa.ipac.caltech.edu/data/SPITZER/docs/dataanalysistools/tools/mopex/}
point-source extraction software for photometry of sources in crowded
fields.  We show spectral energy distributions (SEDs) for the three
high-$\sigmasfr$ galaxies in Figure~\ref{fig:seds}.

@@

We estimate IR-based star-formation rates for the 25/29 galaxies with
WISE 12 or 22~$\mu$m detections by fitting \citet{cha01} templates to
their 12 and 22~$\mu$m fluxes.  For the 14/29 galaxies with 22~$\mu$m
detections, this yields SFRs that agree with those obtained from the
\citet{ruj12} method based on 24~$\mu$m luminosity with a scatter of
0.05~dex.  We note that several authors have shown that the shape of
the IR SED for star-forming galaxies depends on the surface density of
star formation \citep[e.g.,][]{ruj11,elb11}, with more compact
starbursts having larger total-IR (8--1000~$\mu$m) to mid-IR
(8--24~$\mu$m) ratios.  The extreme $\sigmasfr$ values of our sample
imply large total-IR to mid-IR ratios, characteristic of the most
luminous galaxies in the local universe
\citep[e.g.,][]{rie09}.  If we used the most luminous
local templates for the 8/29 sources with SFRs in the ULIRG regime
($\sfrir>200$~\msun~yr$^{-1}$), we would obtain SFRs that are larger
by 0.5~dex than the values we adopt for this paper.

We also estimate star-formation rates and stellar masses based on
stellar population fits to the $\lrest=0.1$--3~$\mu$m SEDs using the
method of \citet{mou11}.  For galaxies with
$\sfrir>50$~\msun~yr$^{-1}$, there is agreement between these UV-based
SFR estimates and $\sfrir$ with a scatter of 0.32~dex.  For an SMC
dust law, we find a median attenuation of $A_V=0.4$~mag.

@@

\begin{figure*}[!t]
\begin{center}
\includegraphics[angle=0,scale=.8]{sigmasfr.ps} 
\caption{SFR surface densities and stellar masses for the HST--WISE
  sample described in this paper (black circles, with symbol size
  proportional to outflow velocity), along with samples of gas-rich
  mergers (orange squares) and star-forming galaxies (shown with 68\%,
  95\%, and 99.7\% contours).  We mark the empirical threshold for
  launching winds \citep[dotted line,
    $\sigmasfr\approx0.1$~\units;][]{hec02}, the 90th-percentile
  starburst intensity limit from \citet{meu97} (dashed line,
  $\sigmasfr\approx45$~\units), and the Eddington limit from radiation
  pressure on dust grains \citep[solid line,
    $\sigmasfr\approx2000$~\units;][]{mur05,tho05,hop10}.  Our
  HST--WISE sample overlaps with the region characterized by gas-rich
  mergers, and extends to very large SFR surface densities near the
  Eddington limit, suggesting growth that is limited by momentum
  injection from massive stars.}
\label{fig:sigmasfr}
\end{center}
\end{figure*}

@@

We present $\lrest=2500$--5200~\AA\ spectroscopy for three
high-$\sigmasfr$ sources in Figure~\ref{fig:spectra} based on data
from MMT/Blue Channel and SDSS (J1506+5402), Magellan/MAGE
(J1341-0321), and Keck/LRIS and Keck/HIRES (J0905+5759).  The spectra
are dominated by light from the young ($t<50$~Myr) stellar population.
We highlight the \mgii~$\lambda\lambda2796,2803$ absorption lines,
which are used used to measure outflow velocities.  At low spectral
resolution (e.g., the top right panel of Figure~\ref{fig:spectra}) it
is not possible to determine the intrinsic shape of the absorption
line profile and therefore the covering factor of the outflowing gas.
However, the Keck/HIRES spectrum of J0905+5759
($\textnormal{FWHM}\approx8$~\kms) reveals that the gas covers the
entire continuum source near the velocity centroid ($v=-2470$~\kms)
indicating a galaxy-wide outflow.

@@

The compact sizes ($r_e\approx100$~pc) and large star formation rates
($\textnormal{SFR}\approx300$~\msun) for the galaxies described above
implies extremely large star-formation rate surface densities
($\sigmasfr\approx5000$~\units).  To place these galaxies in context,
we plot their $\sigmasfr$ values as a function of stellar mass in
Figure~\ref{fig:sigmasfr}.  We include comparison samples of
$\sim10^5$ star-forming galaxies at $0.5<z<1.5$ from \citet{wuy11} and
a few dozen gas-rich mergers at $z<0.3$ including 32 ULIRGs from
\citet{vei06}, five Lyman break analogs with dominant central objects
from \citet{ove09}, and the local compact starburst Arp 220
\citep{sco97, ken98, rod08}.  We also mark the empirical threshold for
launching winds \citep[$\sigmasfr\approx0.1$~\units,][]{hec02}, the
90th percentile limit for the surface brightness of starbursts
measured using UV, H$\alpha$, far-IR, and radio continuum emission
\citep[$\sigmasfr\approx45$~\units][]{meu97}, and the theoretical
limit for a starburst limited by feedback from radiation pressure
\citep[$\sigmasfr\approx2000$~\units,][]{mur05,tho05,hop10}.  The most
luminous, compact starbursts in our sample exhibit SFR surface
densities that reach the Eddington limit, suggesting that their growth
is being regulated by momentum input from massive stars.

@@

While the SEDs (Figure~\ref{fig:seds}) and optical spectra
(Figure~\ref{fig:spectra}) for our sample indicate that their
bolometric output is dominated by star formation, it is worthwhile to
consider the level of ongoing AGN activity and its effect on our
measurements.  Among the high-$\sigmasfr$ galaxies, the strongest case
for detectable AGN emission can be made for J1506+5402, which has the
most luminous \oiii~$\lambda5007$ line among the 29 galaxies observed
by HST.  This source also has a weak \nev~$\lambda3426$ emission line,
which is associated with AGN activity \citep[e.g.,][]{gil10}.  It was
observed with the Chandra X-ray Observatory (proposal ID 11700896, PI:
C. Tremonti) and had four detected counts, corresponding to a
0.5--8.0~keV X-ray luminosity of $10^{42.7}$~erg~s$^{-1}$ (see P. Sell
et al., in preparation for more details on the 12/29 galaxies with
Chandra observations).  Using the relationship between 2--10~keV X-ray
and 12.3~$\mu$m mid-IR luminosity for AGNs from \citet{gan09}, one
would expect a source with
$L_{\textnormal{\scriptsize{X}}}\approx5\times10^{42}$~erg~s$^{-1}$ to
have
$L_{\textnormal{\scriptsize{MIR}}}\approx7\times10^{42}$~erg~s$^{-1}$,
which is 400$\times$ smaller than observed luminosity for J1506+5402
($L_{\textnormal{\scriptsize{MIR}}}\approx3\times10^{45}$~erg~s$^{-1}$).
Based on its \oiii\ luminosity ($10^{42.1}$~erg~s$^{-1}$, a
significant fraction of which could be attributed to star formation)
and the calibration for type 1 AGNs from \citet{hec05}, one would
expect an intrinsic 2--10~keV luminosity of
$L_{\textnormal{\scriptsize{X}}}\approx5\times10^{43}$~erg~s$^{-1}$,
suggesting absorption by a factor of $\sim10$ in the X-rays.  However,
even if the X-ray attenuation were a factor of 100, which is the
typical value for local Compton-thick AGNs \citep{dia09}, the expected
mid-IR AGN contribution would only be relevant at the $\lesssim$30\%
level.  We conclude that the bolometric output of the galaxies in our
sample is dominated by star and we that our results regarding large
$\sigmasfr$ values are not affected by AGN contamination.

@@

It is worth considering whether the high-velocity outflows we observe
were produced by the compact starburst.  \citet{mur11} argued that
massive star clusters with large gas surface densities are the ideal
launching point for galactic-scale outflows driven by radiation
pressure, and that the outflow velocity should scale with escape
velocity of the most massive star clusters in a galaxy.  The
$\gtrsim1000$~\kms\ outflow velocities we observe correspond
approximately to the escape velocities for a compact stellar
population with the sizes and masses that we've measured.
\begin{eqnarray}\nonumber
v_{esc}&=&\sqrt{2GM_{*}/r} \\
&=& 2100
\left({M_{*}\over10^{11} M_{\odot}}\right)^{1/2} 
\left({r\over200~\textnormal{pc}}\right)^{-1/2}
\textnormal{km~s}^{-1} 
\label{eq:vesc}
\end{eqnarray}
This argument, combined with the fact that we observe galaxies with
significant dust-obscured star formation and $\sigmasfr$ values near
the Eddington limit, indicates that momentum input from massive stars
in the form of radiation pressure is a viable mechanism for launching
these outflows.

@@

It is clear from Figure~\ref{fig:sigmasfr} that the high-$\sigmasfr$
galaxies in our sample constitute a rare population, suggesting that
they represent an unusual or short-lived phase.  Mergers of gas-rich
galaxies are a viable mechanism for producing compact starbursts
\citep[e.g.,][]{mih96}, and such gas-rich major mergers are rare at
$z\sim0.6$ due to the decline in both the merger rate
\citep[e.g.,][]{lot11} and the gas fraction \citep[e.g.,][]{tac10} of
galaxies since $z\sim2$.  Furthermore, the high-$\sigmasfr$ galaxies
are caught in a particular time interval where there is both vigorous
star formation and strong feedback.  The length of this phase may be
set by the gas consumption timescale or the timescale for feedback to
suppress subsequent star formation.  Based on the Kennicutt--Schmidt
(K-S) relation \citep{ken98}, a compact starburst with
$r_e\approx100$~pc, $\textnormal{SFR}\approx300$~\msun, and
$\sigmasfr\approx5000$~\units\ would have a gas surface density of
$\Sigma_{gas}\sim2\times10^{11}$~\msun~kpc$^{-2}$ corresponding to
$M_{gas}\sim5\times10^{9}$~\msun\ inside 100~pc, which would be
consumed on a timescale $\sim30$~Myr.  This scenario could be tested
with CO observations of molecular gas masses and kinematics for these
extreme galaxies.

Considering our full sample of galaxies in Figure~\ref{fig:sigmasfr},
a vertical $\sigmasfr$ sequence is apparent.  This could be explained
as an evolutionary sequence where the high-$\sigmasfr$ galaxies
represent the peak of the starburst where the high-velocity outflows
are launched, while the lower $\sigmasfr$ galaxies represent a
subseqent, post-starburst phase.  It is interesting to note that all
12/29 galaxies above the $\sigmasfr\approx45$~\units\ limit from
\citet{meu97} exhibit outflows (with median centroid velocity
$v=-1500$~\kms), while all 7/29 galaxies without detected outflows
have $\sigmasfr<30$~\units.  If a compact starburst is a requirement
for the production of high-velocity outflows, it may be that the
sources without outflows have not gone through such a phase (see
A. Robaina, et al., in preparation for a discussion of the
relationship between galaxy morphology, outflow velocity, and stellar
population age in this sample).

Finally, it is worthwhile to consider the implications of our results
for models of massive galaxy formation.  Simulaitons of major galaxy
mergers with large gas fractions ($f_{gas}\sim50$\%) can produce
$\textnormal{\mstar}\sim10^{11}$~\msun\ remnants with effective radii
$r_{e}\sim1$~kpc \citep[e.g.,][]{wuy10}, but our sample includes
galaxies of similar mass that are smaller by almost an order of
magnitude.  If these systems do evolve into quiescent elliptical
galaxies, it would be quite challenging for them to grow in size from
$r_e\sim0.1$~kpc to $r_e\sim5$~kpc to reach the local size--mass
relation \citep{she03} in the $t\sim6$~Gyr since $z=0.6$.  For
comparison the compact, quiescent galaxies observed at $z\sim2$
\citep[e.g.,][]{tru07,van08} have $t\sim10$~Gyr to grow by a factor of
$\sim5$.  In this context, it is worth noting that the half-light
radius (ignoring dust attenuation) at rest-frame $V$ band (where our
HST observations probe) can be a factor of 5--10 smaller than the
half-mass radius for a gas-rich merger at final coalescence near the
peak of starburst activity \citep{wuy10}.  It would be worthwhile to
test for such size discrepancies and to pronbe the radial dependence
of the mass-to-light ratio for these galaxies by measuring sizes at
rest-frame near-IR wavelengths.

@@

We have measured large SFR surface densities for galaxies that exhibit
$\gtrsim1000$~\kms\ outflows.  The largest $\sigmasfr$ values are
comparable to the Eddington limit from radiation pressure on dust
grains, and such compact starburst are expected to have substantial
momentum input from both massive stars and supernovae.  There is some
evidence for AGN activity in a fraction of these systems, but the AGN
is generally not relevant in a bolometric sense.  High-velocity
outflows have been previously interpreted as a signpost of AGN
feedback, but we conclude that stellar feedback from a compact
starburst is equally capable of producing such a signature.

>>>>

The central regions of elliptical galaxies are thought to form in
compact starbursts \citep[e.g.,][]{kor09,hop09}.  Feedback associated
with such starbursts can produce outflows driven by thermal energy
from supernova explosions \citep[e.g.,][]{che85}, stellar winds
\citep[e.g.,][]{lei92}, and momentum input from both supernova ram
pressure and radiation pressure on dust grains \citep[e.g.,][]{mur05}.
It has been argued that such feedback imposes a limit on the maximum
star-formation rate (SFR) surface density ($\sigmasfr$) for starbursts
\citep[e.g.,][]{leh96,meu97,mur05,tho05} and the maximum stellar
surface density for elliptical galaxies and star clusters
\citep[e.g.,][]{hop10}.

@@

In this Letter, we present results for a sample of massive galaxies at
$z\sim0.6$ that exhibit $\gtrsim1000$~\kms\ outflows, expanding on the
sample from \citet{tre07}.  We seek to test whether the energetic
outflows in these galaxies could have been driven by feedback from
starbursts with very large SFR surface densities.  Our analysis
combines galaxy sizes obtained with the Hubble Space Telescope (HST)
with SFRs and stellar masses estimated using photometry from the
Wide-field Infrared Survey Explorer (WISE), the Spitzer Space
Telescope, the Sloan Digital Sky Survey (SDSS), and the Galaxy
Evolution Explorer (GALEX).

@@

% We gathered photometry from the All-Sky Release of the Wide-field
% Infrared Survey Explorer \citep[WISE,][]{wri10}, the Seventh Data
% Release of the Sloan Digital Sky Survey \citep[SDSS,][]{aba09}, and
% General Release 6 from the Galaxy Evolution Explorer
% \citep[GALEX,][]{mar05galex}.  
We gathered photometry from the WISE All-Sky Release, the SDSS Seventh
Data Release, and GALEX General Release 6.  We also obtained
$5\times30$~sec dithered exposures at 3.6~$\mu$m and 4.5~$\mu$m for
all sources with the Infrared Array Camera on the Spitzer Space
Telescope as part of General Observer program 60145 (PI: C. Tremonti).
We used the post--basic calibrated data to perform aperture photometry
on all sources and point-source photometry on sources in crowded
fields.  We show spectral energy distributions (SEDs) for the three
high-$\sigmasfr$ galaxies in Figure~\ref{fig:seds}.

@@

We estimate IR-based SFRs for the 25/29 galaxies with WISE 12 or
22~$\mu$m detections by fitting \citet{cha01} templates to their 12
and 22~$\mu$m fluxes.  For the 14/29 galaxies with 22~$\mu$m
detections, this yields SFRs that agree with those obtained from the
\citet{ruj12} method based on 24~$\mu$m luminosity with a scatter of
0.05~dex.  We note that several authors have shown that the shape of
the IR SED for star-forming galaxies depends on the surface density of
star formation \citep[e.g.,][]{ruj11,elb11}, with more compact
starbursts having larger total-IR (8--1000~$\mu$m) to mid-IR
(8--24~$\mu$m) ratios.  The extreme $\sigmasfr$ values of our sample
imply large total-IR to mid-IR ratios, characteristic of the most
luminous galaxies in the local universe \citep[e.g.,][]{rie09}.  If we
used the most luminous local templates for the 8/29 sources with SFRs
in the ULIRG regime ($\sfrir>200$~\msun~yr$^{-1}$), we would obtain
SFRs that are larger by 0.5~dex than the values we adopt for this
paper.

We also estimate SFRs and stellar masses based on stellar population
fits to the $\lrest=0.1$--3~$\mu$m SEDs using the method of
\citet{mou11}.  For galaxies with $\sfrir>50$~\msun~yr$^{-1}$, there
is agreement between these UV-based SFR estimates and $\sfrir$ with a
scatter of 0.32~dex.  For an SMC dust law, we find a median
attenuation of $A_V=0.4$~mag.

@@

\begin{figure*}[!t]
\begin{center}
\includegraphics[angle=0,scale=.8]{sigmasfr.ps} 
\caption{SFR surface densities and stellar masses for the HST--WISE
  sample described in this paper (black circles, with symbol size
  proportional to outflow velocity), along with samples of $z<0.3$
  gas-rich mergers (orange squares) and $z\sim1$ star-forming galaxies
  (shown with 68\%, 95\%, and 99.7\% contours).  We mark the empirical
  threshold for launching winds \citep[dotted line,
    $\sigmasfr\approx0.1$~\units;][]{hec02}, the 90th-percentile
  starburst intensity limit from \citet{meu97} (dashed line,
  $\sigmasfr\approx45$~\units), and the Eddington limit from radiation
  pressure on dust grains \citep[solid line,
    $\sigmasfr\approx2000$~\units;][]{mur05,tho05,hop10}.  Our
  HST--WISE sample overlaps with the region characterized by gas-rich
  mergers, and extends to very large SFR surface densities near the
  Eddington limit, suggesting growth that is limited by momentum
  injection from massive stars.}
\label{fig:sigmasfr}
\end{center}
\end{figure*}

@@

We present $\lrest=2500$--5200~\AA\ spectroscopy for three
high-$\sigmasfr$ sources in Figure~\ref{fig:spectra} based on data
from MMT/Blue Channel and SDSS (J1506+5402), Magellan/MagE
(J1341-0321), and Keck/LRIS and Keck/HIRES (J0905+5759).  The spectra
are dominated by light from the young ($t<50$~Myr) stellar population.
We highlight the \mgii~$\lambda\lambda2796,2803$ absorption lines,
which are used used to measure outflow velocities.  At low spectral
resolution (e.g., the top right panel of Figure~\ref{fig:spectra}) it
is not possible to determine the intrinsic shape of the absorption
line profile and therefore the covering factor of the outflowing gas.
However, the Keck/HIRES spectrum of J0905+5759
($\textnormal{FWHM}\approx8$~\kms) reveals that the gas covers the
entire continuum source near the velocity centroid ($v=-2470$~\kms)
indicating a galaxy-wide outflow.

@@

The compact sizes ($r_e\approx100$~pc) and large SFRs
($\textnormal{SFR}\approx300$~\msun) for the galaxies described above
implies extremely large SFR surface densities
($\sigmasfr\approx5000$~\units).  To place these galaxies in context,
we plot their $\sigmasfr$ values as a function of stellar mass in
Figure~\ref{fig:sigmasfr}.  We include comparison samples of
$\sim10^5$ star-forming galaxies at $0.5<z<1.5$ from \citet{wuy11} and
a few dozen gas-rich mergers at $z<0.3$ including 32 ULIRGs from
\citet{vei06}, five Lyman break analogs with dominant central objects
from \citet{ove09}, and the local compact starburst Arp 220
\citep{sco97, ken98, rod08}.  We also mark the empirical threshold for
launching winds \citep[$\sigmasfr\approx0.1$~\units,][]{hec02}, the
90th percentile limit for the surface brightness of starbursts
measured using UV, H$\alpha$, far-IR, and radio continuum emission
\citep[$\sigmasfr\approx45$~\units][]{meu97}, and the theoretical
limit for a starburst limited by feedback from radiation pressure
\citep[$\sigmasfr\approx2000$~\units,][]{mur05,tho05,hop10}.  The most
luminous, compact starbursts in our sample exhibit SFR surface
densities that reach the Eddington limit, suggesting that their growth
is being regulated by momentum input from massive stars.

@@

While the SEDs (Figure~\ref{fig:seds}) and optical spectra
(Figure~\ref{fig:spectra}) for our sample indicate that their
bolometric output is dominated by star formation, it is worthwhile to
consider the level of ongoing AGN activity and its effect on our
measurements.  Among the high-$\sigmasfr$ galaxies, the strongest case
for detectable AGN emission can be made for J1506+5402, which has the
most luminous \oiii~$\lambda5007$ line among the 29 galaxies observed
by HST.  This source also has a weak \nev~$\lambda3426$ emission line,
which is associated with AGN activity \citep[e.g.,][]{gil10}.  It was
observed with the Chandra X-ray Observatory (proposal ID 11700896, PI:
C. Tremonti) and had four detected counts, corresponding to a
0.5--8.0~keV X-ray luminosity of $10^{42.7}$~erg~s$^{-1}$ (see P. Sell
et al., in preparation for more details on the 12/29 galaxies with
Chandra observations).  Using the relationship between 2--10~keV X-ray
and 12.3~$\mu$m mid-IR luminosity for AGNs from \citet{gan09}, one
would expect a source with
$L_{\textnormal{\scriptsize{X}}}\approx5\times10^{42}$~erg~s$^{-1}$ to
have
$L_{\textnormal{\scriptsize{MIR}}}\approx7\times10^{42}$~erg~s$^{-1}$,
which is 400$\times$ smaller than observed luminosity for J1506+5402
($L_{\textnormal{\scriptsize{MIR}}}\approx3\times10^{45}$~erg~s$^{-1}$).
Based on its \oiii\ luminosity ($10^{42.1}$~erg~s$^{-1}$, a
significant fraction of which could be attributed to star formation)
and the calibration for type 1 AGNs from \citet{hec05}, one would
expect an intrinsic 2--10~keV luminosity of
$L_{\textnormal{\scriptsize{X}}}\approx5\times10^{43}$~erg~s$^{-1}$,
suggesting absorption by a factor of $\sim10$ in the X-rays.  However,
even if the X-ray attenuation were a factor of 100, which is the
typical value for local Compton-thick AGNs \citep{dia09}, the expected
mid-IR AGN contribution would only be relevant at the $\lesssim$30\%
level.  We conclude that the bolometric output of the galaxies in our
sample is dominated by star formation and we that our results
regarding large $\sigmasfr$ values are not affected by AGN
contamination.

@@

It is worth considering whether the high-velocity outflows we observe
could have been produced by the compact starburst.  \citet{mur11}
argued that massive star clusters with large gas surface densities are
the ideal launching point for galactic-scale outflows driven by
radiation pressure, and that the outflow velocity should scale with
escape velocity of the most massive star clusters in a galaxy.  The
$\gtrsim1000$~\kms\ outflow velocities we observe correspond
approximately to the escape velocities for a compact stellar
population with the sizes and masses that we've measured.
\begin{eqnarray}\nonumber
v_{esc}&=&\sqrt{2GM_{*}/r} \\
&=& 2100
\left({M_{*}\over10^{11} M_{\odot}}\right)^{1/2} 
\left({r\over200~\textnormal{pc}}\right)^{-1/2}
\textnormal{km~s}^{-1} 
\label{eq:vesc}
\end{eqnarray}
This argument, combined with the fact that we observe galaxies with
significant dust-obscured star formation and $\sigmasfr$ values near
the Eddington limit, indicates that momentum input from massive stars
in the form of radiation pressure is a viable mechanism for launching
these outflows.

@@

It is clear from Figure~\ref{fig:sigmasfr} that the high-$\sigmasfr$
galaxies in our sample constitute a rare population, suggesting that
they represent an unusual or short-lived phase.  Mergers of gas-rich
galaxies are a viable mechanism for producing compact starbursts
\citep[e.g.,][]{mih96}, and such gas-rich major mergers are rare at
$z\sim0.6$ due to the decline in both the merger rate
\citep[e.g.,][]{lot11} and the gas fraction \citep[e.g.,][]{tac10} of
galaxies since $z\sim2$.  Furthermore, the high-$\sigmasfr$ galaxies
in our sample are caught in a particular time interval where there is
both vigorous star formation and strong feedback.  The length of this
phase may be set by the gas consumption timescale or the timescale for
feedback to suppress subsequent star formation.  Based on the
Kennicutt--Schmidt (K-S) relation \citep{ken98}, a compact starburst
with $r_e\approx100$~pc, $\textnormal{SFR}\approx300$~\msun, and
$\sigmasfr\approx5000$~\units\ would have a gas surface density of
$\Sigma_{gas}\sim2\times10^{11}$~\msun~kpc$^{-2}$ corresponding to
$M_{gas}\sim5\times10^{9}$~\msun\ inside 100~pc, which would be
consumed on a timescale $\sim30$~Myr.  This scenario could be tested
with CO observations of molecular gas masses and kinematics for these
extreme galaxies.

Considering our full sample of galaxies in Figure~\ref{fig:sigmasfr},
a vertical $\sigmasfr$ sequence is apparent.  This could be explained
as an evolutionary sequence where the high-$\sigmasfr$ galaxies
represent the peak of the starburst when the high-velocity outflows
are launched, while the lower $\sigmasfr$ galaxies represent a
subsequent, post-starburst phase.  It is interesting to note that all
12/29 galaxies above the $\sigmasfr\approx45$~\units\ limit from
\citet{meu97} exhibit outflows (with median centroid velocity
$v=-1500$~\kms), while all 7/29 galaxies without detected outflows
have $\sigmasfr<30$~\units.  If a compact starburst is a requirement
for the production of high-velocity outflows, it may be that the
sources without outflows have not gone through such a phase (see
A. Robaina, et al., in preparation for a discussion of the
relationship between galaxy morphology, outflow velocity, and stellar
population age in this sample).

Finally, it is worthwhile to consider the implications of our results
for models of massive galaxy formation.  Simulations of major galaxy
mergers with large gas fractions ($f_{gas}\sim50$\%) can produce
$\textnormal{\mstar}\sim10^{11}$~\msun\ remnants with effective radii
$r_{e}\sim1$~kpc \citep[e.g.,][]{wuy10}, but our sample includes
galaxies of similar mass that are smaller by almost an order of
magnitude.  If these systems do evolve into quiescent elliptical
galaxies, it would be extremely challenging for them to grow in size
from $r_e\sim0.1$~kpc to $r_e\sim5$~kpc to reach the local size--mass
relation \citep{she03} in the $t\sim6$~Gyr since $z=0.6$.  For
comparison the compact, quiescent galaxies observed at $z\sim2$
\citep[e.g.,][]{tru07,van08} have $t\sim10$~Gyr to grow by a factor of
$\sim5$.  In this context, it is worth noting that the half-light
radius (ignoring dust attenuation) at rest-frame $V$ band (where our
HST observations probe) can be a factor of 5--10 smaller than the
half-mass radius for a gas-rich merger at final coalescence near the
peak of starburst activity \citep{wuy10}.  It would be worthwhile to
test for such size discrepancies and to probe the radial dependence of
the mass-to-light ratio for these galaxies by measuring sizes at
rest-frame near-IR wavelengths.

@@

We have measured large SFR surface densities for galaxies that exhibit
$\gtrsim1000$~\kms\ outflows.  The largest $\sigmasfr$ values are
comparable to the Eddington limit from radiation pressure on dust
grains, and such compact starburst are expected to have substantial
momentum input from massive stars and supernovae.  There is some
evidence for AGN activity in a fraction of these systems, but the AGN
is generally not relevant in a bolometric sense.  While high-velocity
outflows have been previously interpreted as a signpost of AGN
feedback, we conclude that stellar feedback from a compact starburst
is equally capable of producing such a signature.
